### 2013-02-22 doc/xps2013/proposal.tex
minor edits

\vspace{1ex} 
\noindent\textbf{Year 3:} In the third and final year of this
effort (supported by this grant), we will finish the networking layer rewrite.
We will also have constructed a test-bed system for the SSD-only data
structure research. Grappa will be ported to this new style of memory
hierarchy, which we envision will largely be a tuning and not rewriting
effort. Finally we will complete a detailed system evaluation using our
benchmark suite.

@@

Irregular applications are increasing in importance to the computing industry.
For a long time they have underpinned computations of interest to the national
security arm of the federal government. But now, they underpin such questions
as ad placement in social networks, and analysis of complex data-sets in
medicine and science. The defining characteristic of these applications is
poor locality and massive available parallelism. Our effort is focused on
making these applications perform well and scale on commodity hardware. The
key idea that makes this work is to rely on massive concurrency to tolerate
memory latency, instead of relying on locality (which doesn't exist for these
applications). This core idea underpinned the XMT~\cite{feo:xmt} system, a
fully-custom hardware system. We plan to explore that idea in software only,
running on commodity processors. And unlike the XMT, however, we must also
contend with commodity networking hardware, and here we again rely on
threading to provide a sufficient number of requests that we can buffer them
prior to dispatching on the network, to overcome bandwidth limitations when
operating with small messages.

>>>>

\vspace{1ex} 
\noindent\textbf{Year 3:} In the third and final year of this effort (supported by this grant), we will finish the networking layer rewrite. We will also have constructed a test-bed system for the SSD-only data structure research. Grappa will be ported to this new style of memory hierarchy.  Finally we will complete a detailed system evaluation using our benchmark suite.

@@

Irregular applications are increasing in importance to the computing industry. For a long time they have underpinned computations of interest to national security. But now, they underpin such questions as ad placement in social networks, and analysis of complex data-sets in medicine and science. The defining characteristic of these applications is poor locality and massive available parallelism. Our effort is focused on making these applications perform well and scale on commodity hardware. The key idea that makes this work is to rely on massive concurrency to tolerate memory latency, instead of relying on locality (which doesn't exist for these applications). This core idea underpinned the XMT~\cite{feo:xmt} system, a fully-custom hardware system. We plan to explore that idea in software only, running on commodity processors. And unlike the XMT, however, we must also contend with commodity networking hardware, and here we again rely on threading to provide a sufficient number of requests that we can buffer them prior to dispatching on the network, to overcome bandwidth limitations when operating with small messages.

### 2013-02-22 doc/xps2013/summary.tex

Our preliminary effort has shown a workable proof of concept of this approach: a commodity cluster can out do the XMT at its own game, by emulating its key features in software. What is needed now is additional effort to improve the underlying performance of this runtime and make it usable to a broad base of users. We will be focusing our efforts in the next three years on the key performance component, the networking layer, and the key programmer efficiency linchpin, high level language support. In addition we need a useful mechanism for handling data-sets in the petabyte range, and here we imagine building a cluster of machines that relies primarily on SSD's and not DRAM as main memory data-set storage.

If successful, the Grappa runtime will have a transformative impact on the computing industry. Irregular applications are hugely important, and if off the shelf hardware can be efficient with them a new class of questions can be asked of computing systems. Additionally our benchmark suite development will further drive innovation among researchers in this space.

>>>>

Our preliminary effort has shown a workable proof of concept of this approach: a commodity cluster can out do the XMT at its own game, by emulating its key features in software. What is needed now is additional effort to improve the underlying performance of this runtime and make it usable to a broad base of users. We will be focusing our efforts in the next three years on the key performance component, the networking layer, and the key programmer efficiency linchpin, high level language support. In addition we need a useful mechanism for handling data-sets in the petabyte range, and here we imagine building a cluster of machines that relies primarily on SSD's and not DRAM as main  data-set storage.

If successful, the Grappa runtime will have a transformative impact on the computing industry. Irregular applications are hugely important, and if off-the-shelf hardware can be efficient with them a new class of questions can be asked of computing systems. Additionally our benchmark suite development will further drive innovation among researchers in this space.
